### 2016-05-25 paper.Rmd
yo-yo

---
title: "Automated optimization of a course of actions in a stochastic environment"
author: "Thomas Vicente"
output: pdf_document
fontsize: 12pt
header-includes: \usepackage{bm}
                 \usepackage{bbm}
                 \usepackage{algorithmic}
                 \usepackage{algorithm}
---

#General principle

\newcommand{\IND}[1][1]{\STATE\hspace{#1\algorithmicindent}}
\begin{algorithm*}
\caption{General version}
\begin{algorithmic}
\STATE$\textbf{initialize:}$
\IND$f() \text{: regressor function}$
\IND$x_t \text{: current stochastic environment}$
\IND$\Omega_t(x_t) \text{: set of possible actions}$
\IND$s(x_t) \text{: signal activated or not by the stochastic environment, initialized to 0}$
\IND$W^*=W_0\text{: random weights for the regressor}$
\IND$X\text{: empty training feature matrix}$
\IND$y\text{: empty training target vector}$
\IND$A\text{: empty feature matrix for simulated actions}$
\IND$\bm{\gamma} \in [0,1] \text{ discount rate}$
\IND$\Delta = 0 \text{: number of actions}$
\STATE$\textbf{while } \text{the learning process is on:}$
\IND$\text{record }x_t$
\IND$A \text{ = empty matrix}$
\IND$\textbf{for } a' \text{ simulated action in } \Omega_t(x_t)$
\IND[2]$A.append(a')$
\IND$a^*=\arg\max_{a' \in A} f(W^*,A)$
\IND$X.append(a^*)$
\IND$\textbf{if } s(a^*) \text{ != } 0 \text{ :}$
\IND[2]$\textbf{for } a \text{ in } \Delta \text{:}$
\IND[3]$value = \gamma^{a-1}s(a^*)$
\IND[3]$y.append(value)$
\IND[3]$W^* := \arg\min_{W}(l(f(W,X),y))$
\IND[3]$\Delta = 0$
\IND$\textbf{else} \text{:}$
\IND[2]$\Delta \text{ += } 1$
\IND[2]$\text{continue}$
\end{algorithmic}
\end{algorithm*}

The algorithm assumes two underlying function:

- The minimization process is a tedious part of the algorithm. When dealing with unstructured features, we might want to use a neural network-type regressor. In that case, $W$, the set of the hidden layers' weights is initialized with the values of the previous iterations. The optimal $W^*$ should converge as $X$ grows if there are patterns in the stochastic processes.

- The signal function assumes the existence of one or multiple sensors, or the manual assignement of a value. It is conditional on the current environment and takes positive value if a "gain" is sensed, and negative value if a "pain" is sensed.

###Remark on the function of the regressor's role
The regressor's task is not to predict accurately, even though it is closely related, but rather to give the "real value" of an action.

#Analysis of why RL+NN converge, why it works
It needs to have a good balance between gains and pains

#Application

###Remark on the function of the regressor's role
Two intermediary movements can ultimately lead to both winning and losing. Such states will get a non-tendencial value as $X$ grows, which is a good thing.

##Analysis of game
Do some graph based on data.txt's target variable, observe time effects, wins evolution